\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-02-03 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

\begin{figure}[h]
\makebox[\textwidth]{%
\includegraphics[width=0.49\textwidth]{./Pics/2b2_P1DGP2_perm}%
\hfill    
\includegraphics[width=0.49\textwidth]{./Pics/2b2_P1DGP2_mesh}%
}\\[0.5cm]% If you want some vertical space
\makebox[\textwidth]{%
\includegraphics[width=0.75\textwidth]{./Pics/2b2_P1DGP2_plot}%
}%
\caption{ Caption 1}
\label{fig:1}
\end{figure}

@@

\begin{figure}[h]

\begin{center}
\centering
\includegraphics[width=0.75\linewidth, height=5cm]{./Pics/P1DG-P2_element_type} 
\caption{Caption 2}
\label{fig:2}
\end{center}

\begin{center}
\centering
\includegraphics[width=0.8\linewidth, height=7cm]{./Pics/4r_po_adapt_fine_125_mesh} 
\caption{Caption 3}
\label{fig:3}
\end{center}

\begin{center}
\centering
\includegraphics[width=0.8\linewidth, height=7cm]{./Pics/4r_po_adapt_fine_100_mesh}
\caption{Caption 4}
\label{fig:4}
\end{center}

%\caption{Caption for this figure with two images}
\label{fig:image2}
\end{figure}
\pagebreak
\clearpage

@@

Starting from the simplest case of simulations as describe in fig. \ref{fig:1}, there is a simple square shape formation consisted of 2 values of different (K) permeabilities. We inject fluid 1 across the left hand side of the formation and we are trying to displace the existing fluid 2. The front start to collapse and one main finger starts to appear.

The behaviour as well as the prediction of the two different phase volume fractions can be qualitatively compared with bibliographic results. The two phase volume volume fractions have the same but opposite behaviour across the length of the formation. As we inject fluid 1, at the first half of the formation the velocity magnitude is increasing while pressure drops gradually. This behaviour changes at the middle of the formation where there is the transition from one permeability zone to the other and the front has already collapsed forming the finger.

From that point till the end of the formation, the velocity magnitute of the injected fluid remains the same having the minimum value since there is no more fluid 2 to be displaced. There is a sharp pressure drop which is also an indication that the fluid 2 cannot be pushed any further since at the same time, as can been seen from the graph plot at fig. \ref{fig:1} the phase volume fraction is constant. (\textcolor{blue}{KC: i think we can rephrase it better here.})

Later, we introduce two versions of a second set of simulations consisted of the same number of regions (4) but with different formations and phase ratios as can be described in fig. \ref{fig:2} to fig. \ref{fig:subim7}.

The initial and boundary condition regarding the inlet and outlet are the same. The purpose is to expand this technique and introduce it into more complex formations so as to test it and develop a more stable and solid approach in real time and scale problems.

The main body of the formation (blue area) is where fluid 2 exists fig. \ref{fig:3}. Above the blue area there are two layers - formations of different absolute permeabilities while in the main body there is a also a square shaped red coloured area where the two fluids exist in 50-50 phase ratio. All those ratios as well as the values of the permeabilities can be changes any time according to the data that are provided.

Fluid 1 is again injected across the left hand side of the formation. In this case the value of mobility or \textit{viscosity ratio} is 10. Later on mobility ratios of different values will be introduced and compared. As the flow is developed the initial uniform front collapses start forming fingers, areas where the flow is developed faster than the rest of the main body of the flow. The new unstructured and adaptive mesh with the $P_{n}DG-P_{n+1}$ element type is used in order to capture the flow development. In fig.\ref{fig:2} and fig.\ref{fig:3}) the mesh deforms on time and on space and that allows the user to have an initial good qualitative approach regarding the development of the multiphase flow. In fig. \ref{fig:subim6} and fig. \ref{fig:subim7} the patterns of the multiphase flows for the phase 1 and phase 2 are described respectively.

@@

Comparing the three plots we can see three different velocities to be developed which means that numerically the initial uniform front collapses. This behaviour can be described on the following chapter, where the explanation of viscous instabilities is given.

The first example is a simple comparison between two cases using fixed mesh and mobility ratios of $1$ and $10$ respectively. The aspect ratio of the $2D$ domain is 2 and it favors viscous fingering effects while temperature is not taken into account (assume isothermal environment) and gravity effects are negligible. Fluid $1$ is initially injected from the left hand side with velocity $u_{inj}=1$. The domain is fully saturated with fluid $2$. At the outlet (right hand side) boundary pressure levels are set to zero while no slip boundary conditions are implied at the upper and lower part of the formation. Fixed time step sizes of $1 \times 10^{-3}$ for both the fixed meshes and the CFL number for this set of simulations is 5. The $P_{1}DGP_{2}$ element type with piecewise linear description of veleocity within each element and quadratic pressure representation as part of the OCVFEM method along with the discontinuous Galerkin method are implied and composing the mesh. Saturation is located at the pressure nodes. Five different regions of 5 distinct absolute permeabilities (K) are introduced (fig.\ref{fig:5regions_domain}). The red square region is saturated with a $50\%-50\%$ mixture of both phase $1$ and $2$.

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.5\textwidth]{./Pics1/mr1_fixed/permeability.png}
\caption{2D domain that consist of five regions with 5 different absolute permeabilities as these are represented with different colors.}
\label{fig:5regions_domain}
\end{center}
\end{figure}

>>>>

@@

@@

Starting from a simple case as this is described in fig. \ref{fig:1}, there is a simple square shape formation consisted of 2 values of different (K) permeabilities. The domain is fully saturated with fluid $2$. We inject fluid $1$ across the left hand side of the formation and we are trying to displace the existing fluid $2$. The front start to collapse at the first half of the lenght D (D:distance) of the domain ($x= \frac{1}{2} D)$ and a main finger starts to appear.

\begin{figure}[h]
\makebox[\textwidth]{%
\includegraphics[width=0.39\textwidth]{./Pics/2b2_P1DGP2_perm}%
\hfill    
\includegraphics[width=0.39\textwidth]{./Pics/2b2_P1DGP2_mesh}%
}\\[0.5cm]% If you want some vertical space
\makebox[\textwidth]{%
\includegraphics[width=0.5\textwidth]{./Pics/2b2_P1DGP2_plot}%
}%
\caption{A simple test-case of a square region with four subregions. The permeability contrast is constant. The mesh is unstructured but fixed. After the injection of fluid $1$ at a velocity of u = 1, the velocity magnitude drops to zero. The main finger starts to dies out during the transition from one permeability zone to another and as it approaches closer to the right boundary. The mechanism that still drives the phenomenon is the pressure difference and the tip-splitting pattern can be more profound. }
\label{fig:square}
\end{figure}

In this test-case above the two phase volume volume fractions have the same but opposite behaviour across the length of the formation. From that point till the end of the formation, the velocity magnitute of the injected fluid remains the same having the minimum value since there is no more fluid 2 to be displaced. There is a sharp pressure drop which is also an indication that the fluid 2 cannot be pushed any further since at the same time, as can been seen from the graph plot at fig. \ref{fig:square}, the phase volume fraction is constant.

The initial and boundary condition regarding the inlet and outlet are the same. One more test-case (fig.\ref{fig:testcase}), to verify the finger formation in a simple but same set-up is also illustrated.

\begin{figure}[h]
\begin{center}
\subfigure[MR1 fixed mesh start aspect ratio 2]{%
\label{fig:}
\includegraphics[width=0.45\textwidth]{./Pics1/2b2_wi_fine/2b2_whole_in_fine_250_1.png}
}%
\subfigure[MR1 fixed mesh end aspect ratio 2]{%
\label{fig:}
\includegraphics[width=0.45\textwidth]{./Pics1/2b2_wi_fine/2b2_whole_in_fine_3000_1.png}
}\\
\end{center}
\caption{In this tes-case fixed and unstructured mesh is used. The aspect ratio is 2, which allows the finger to develop faster and make the tip-splitting stage more obvious. Simulation time step increases from left (a) to right (b) and from t=250(start) to t=3000(end).}
\label{fig:testcase}
\end{figure}

 The purpose is to expand this technique and introduce it into more complex formations so as to test it and develop a more stable and solid approach in real time and scale problems. A more detailed description of the element type and the formation that will be used in the next set of test cases follows.

%{./Pics/P1DG-P2_element_type} 
%{./Pics/4r_po_adapt_fine_125_mesh} 
%{./Pics/4r_po_adapt_fine_100_mesh}

\begin{figure}[h]
\begin{center}
\subfigure[The $P_{1}DGP_{2}$ element type]{%
\label{fig:}
\includegraphics[width=0.45\textwidth]{./Pics/P1DG-P2_element_type}
}%
\subfigure[The map of permeabilities]{%
\label{fig:}
\includegraphics[width=0.45\textwidth]{./Pics/4r_po_adapt_fine_125_mesh}
}\\
\end{center}
\caption{(a) represents the element type of $P_{1}DGP_{2}$ as well as where each of the parameters is located and calculated on the element type. (b) describes the permeability map as this is introduced throught a python script during the set-up of simulations. In general, the main body of the formation is the blue area that is fully saturated with fluid $2$. Above the main (blue colored) area there are two layers - formations of different absolute permeabilities while in the main body there is a also a square shaped red coloured area where the two fluids exist in 50-50 phase ratio. All those ratios as well as the values of the permeabilities can be changes any time according to the data that are provided.   }
\label{fig:elem_type_perm_map}
\end{figure}

Fluid 1 is again injected across the left hand side of the formation. Mobility ratios of different values will be introduced and compared. As the flow is developed the initial uniform front collapses start forming fingers, areas where the flow is developed faster than the rest of the main body of the flow. The new unstructured and adaptive mesh with the $P_{n}DG-P_{n+1}$ element type is used in order to capture the flow development. In fig. \ref{fig:simple_case} the simple case of a domain with uniform permeability is displayed.

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\textwidth]{./Pics1/phase_vol_frac_unif_perm.pdf}
\caption{This is the simple case where the initially piston shape front is collapsing and starts forming fingers, in a domain with uniform permeability.}
\label{fig:simple_case}
\end{center}
\end{figure}

@@

Comparing the three plots we can see three different velocities to be developed which means that numerically the initial uniform front collapses. This behaviour can be described on the following chapter, where the explanation of viscous instabilities is given.         
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The first example is a simple comparison between two cases using fixed mesh and mobility ratios of $1$ and $10$ respectively. The aspect ratio of the $2D$ domain is 2 and it favors viscous fingering effects while temperature is not taken into account (assume isothermal environment) and gravity effects are negligible. Fluid $1$ is initially injected from the left hand side with velocity $u_{inj}=1$. The domain is fully saturated with fluid $2$. At the outlet (right hand side) boundary pressure levels are set to zero while no slip boundary conditions are implied at the upper and lower part of the formation. Fixed time step sizes of $1 \times 10^{-3}$ for both the fixed meshes and the CFL number for this set of simulations is 5. The $P_{1}DGP_{2}$ element type with piecewise linear description of veleocity within each element and quadratic pressure representation as part of the OCVFEM method along with the discontinuous Galerkin method are implied and composing the mesh. Saturation is located at the pressure nodes. Five different regions of 5 distinct absolute permeabilities (K) are introduced (fig.\ref{fig:5regions_domain}). The red square region is saturated with a $50\%-50\%$ mixture of both phase $1$ and $2$.

@@
